Deep Convolutional networks for large scale image classification - A Survey

### 2015-01-27 paper/main.tex
draft v2

Early in that stage, method like the convolutional network is limited by the  availability of labeled  data and computing infrastructure.  However, in the recent years, the development of  High performance computing architecture such as general purpose graphical processing units (GPGPU )  accelerated the research in this field. Large scale image  dataset such as ImageNet\cite{imagenet}  with  millions of labeled samples is also accessible to the research community. This changes in data and computing, put back the convolution network with million of parameters in track.

\section{Multi-stage Hubel-Wiesel architecture}
In 1962, Hubel DH and Wiesel TN\cite{Hubel1962},\cite{Hubel1965a} was studied visual cortex of anesthetized cats  with   spots of white light of various shapes. They classified the  cells in the visual system into  simple, complex and hypercomplex. Simple cells are influenced  by the arrangement of  excitatory and inhibitory regions of the receptive field, and  position of the stimulus is important. This cell receives input from cells of the lateral geniculate nucleus(LGN), which is connected to the retina. However, the complex cells will responds to  a properly  oriented stimulus regardless of the cell position in the receptive field. Complex cells are activated by edge, dark bar, slit and mixed stimuli. Hypercomplex cells are activated by edge,single-stopped(corner),double-stopped (tongue),slit(double-stopped)and dark bar(double-stopped).

\par Hubel DH and Wiesel TN \cite{Hubel1965a} presented a functional hierarchical structure of the visual cortex. According to their model, visual perception cells are in the  order, simple$\Longrightarrow$ complex$\Longrightarrow$ lower order hypercomplex $\Longrightarrow$ higher  order hypercomplex. Activation of  a lower stage is  influenced by  the position of the input patterns, and  higher stages are  position-invariant. There are several  contradictory to this structure, but no one   completely deny this hierarchical model.
\par Inspired by this work, Fukushima, K \cite{Stark1980} proposed a neural network model for pattern recognition called neocognitron.
In neocognitron , cells are arranged in  a number of  cascaded structure. Each  structure $U$ include a  simple cell layer $U_s$ and a complex cell layer $U_c$. This network is not affected by change in position or small distortion in the shape of patterns.  It is also capable of doing self-organization based on an unsupervised competitive learning algorithm\cite{Fukushima1982} in the first two layers and classification based on  supervised learning in the output layer.
\begin{figure}[ht]
 \centering
 \includegraphics[width=9cm]{Figures/neocognitron.eps}
\caption{Neocognitron\cite{Stark1980}}
\label{neo}
\end{figure}
\section{Convolutional Networks}
\par Neocognitron was improved by  Yann LeCun \cite{lecun-86},\cite{lecun-89d}, \cite{lecun-89e},\cite{lecun-90c}, \cite{lecun-90e} using backpropagation algorithm\cite{BRYSON1963} to train the entire system.
It uses  local receptive fields, share weights and  sub-sampling to achieve  shift, position and distortion invariance. A typical Convolutional Network called  LeNet-5 was proposed by   Yann LeCun et al..\cite{LeCun1998}  for document recognition. Using local receptive fields network can extract elementary visual features such as edges, end point, and corners. This features will combine to obtain high order features in the following layers.  Elementary feature detectors with identical weights can be useful in different parts of the image. So the units with  the same set of weights are arranged in plane and output from  the units of a plane is called  a feature map. Units in  a feature map perform the same operation on different parts of the same image. A convolution layer is composed of the set of feature maps with differently weighed units. In the implementation, a unit in the feature map scans the image and store the states in the feature map. This operation is equivalent to convolution with a kernel composed of a  set of weights and image.

\par
A typical convolutional  network composed of multiple stages with a filter bank layer, a non-linearity layer and a feature pooling layer \cite{lecun2010convolutional} followed by a classification network.\\
\begin{figure}[ht]
 \includegraphics[width=9.5cm]{Figures/convent.eps}
\caption{A typical ConvNet architecture \cite{lecun2010convolutional}}
\label{net}
\end{figure}
\emph{Filter Bank Layer:} This layer computes $y_{j}$ the convolution between a input feature map   $x_i$ and trainable filter kernel $k_{ij}$ 
. ie. $y_j=b_j+\sum_i {k_{ij}*x_i}$. Where $b_j$ is a trainable bias , $i$ and $j$ are  array indices.\\
\emph{Non-Linearity Layer:} This layer applies a non-linearity function such as $tanh(x)$ or $(1+e^{-x})^{-1}$ to unit output. But to reduce training time with gradient decent, new implementations uses the function $max(0,x)$. Units with this non-linearity is called Rectified Linear Units (ReLUs)\cite{Nair2010}.\\
\emph{Feature Pooling Layer:} It  reduces the dimension of  feature map by applying the techniques like averaging or max-pooling.
\begin{figure}[ht]
 \centering
 \includegraphics[width=8cm]{Figures/conv.eps}
\caption{Output of  a randomly initialized   convolution filter. Input image courtesy:dcsku.org}
\label{conv}
\end{figure}

%2012

\section {Deep convolutional networks}
In last few years convolutional networks shows a significant performance improvement in many small scale image classification on data sets  such as MNIST\cite{Ciresan:2012g},CIFAR-10,CIFAR-100,SVHN\cite{lee2014deeply},STL-10 \cite{deepfwd}. One of the state-of-the-art error rate is from Krizhevsky, et al.\cite{Krizhevsky2012a}. Their network has 60 million parameters and 650,000 neurons, consists of five convolutional layers followed by max-pooling layers, and three fully-connected layers. Data set used in this experiment was a subset of ImageNet dataset, used in the competition ImageNet Large-Scale Visual Recognition Challenge (ILSVRC)\cite{imagenet} and reported an error rate of 15.3\%. This data set includes 1.2 million images that contain 1,000 categories.

@@

\subsection{Network In Network }
Inspired by the work of Ian J. Goodfellow et al..\cite{Goodfellow2013} on max out networks,  Min Lin et al. \cite{Lin2013} introduced a micro-network in each convolution layer so that it will compute more abstract features.This network gave a state-of-the-art performance in  ILSVRC 2013 competition with an error rate of 12.95\%.They used NVIDIA TITAN GPU to train the network.

@@

Matthew D. Zeiler and Rob Fergus\cite{Zeiler2013} presented a method to visualize the function of intermediate feature layers of convolutional networks and used as a diagnostic tool to improve  the model proposed by Krizhevsky et al. \cite{Krizhevsky2012a}. This method helped  them  to understand  the activation in the feature maps with respect to the input patterns. It shows Krizhevsky et al..'s architecture does not have  enough mid frequency coverage in the first layer filters and aliasing artifacts caused by  large stride in the first layer convolutions. Authors solved this problems by decreasing filter size to $7\times7$ and reducing stride to 2.This implementation won the  ILSVRC 2013 competition with an error rate of 11.74\%.

@@

\subsection{Spatial pyramid pooling in Deep Convolutional Networks}
Instead of using fixed input size in convolutional networks, Kaiming He et al. \cite{He2014} suggested  the use of a spooling strategy called  spatial pyramid pooling(SPP)\cite{Grauman2005}\cite{1641019} to avoid cropping or warping of  images. It introduced a new layer on top of the convolution layer and perform aggregation  based on Bag-of-Words (BoW) model \cite{Sivic2003}. However, the classical backpropagation training methods expect layers to have a fixed   size. To overcome this problem author implemented two fixed size networks with shared parameters and switched the network on alternate epochs. This network is trained using a single GeForce GTX Titan GPU with a starting  learning rate of 0.01 and achieved  a less  error rate of 8.06\% on  ILSVRC 2014 data set. 
\par 
This implementation improves the performance of baseline architectures including ZF-5\cite{Zeiler2013} , Convnet \cite{Krizhevsky2012a} and Overfeat-5/7 \cite{Sermanet2013}. Their study shows; accuracy of convolutional networks will improve on multi-size training, multi-level pooling, and full-image representations. 
%

>>>>

Early in that stage, method like the convolutional network is limited by the  availability of labeled  data and computing infrastructure. Researchers are trying to overcome  this problem by many other non-parametric models such as SVM and KNN. But this methods couldn't give a high accuracy result  on any of  large-scale classification problem  and limited by the preprocessing technique.  However, in the recent years, the development of  High performance computing architecture such as general purpose graphical processing units (GPGPU )  accelerated the research in this field. Large scale image  dataset such as ImageNet\cite{imagenet}  with  millions of labeled samples is also accessible to the research community. This changes in data and computing, put back the convolution network with million of parameters in track.

\section{Multi-stage Hubel-Wiesel architecture}
In 1962, Hubel DH and Wiesel TN\cite{Hubel1962},\cite{Hubel1965a} was studied visual cortex of anesthetized cats  with   spots of white light of various shapes. Cells in the visual system are classified    into  simple, complex and hypercomplex. Simple cells are influenced  by the arrangement of  excitatory and inhibitory regions of the receptive field, and  position of the stimulus is important. This cell receives input from cells of the lateral geniculate nucleus(LGN), which is connected to the retina. However, the complex cells will responds to  an appropriately  oriented stimulus regardless of the cell position in the receptive field. Complex cells are activated by edge, dark bar, slit and mixed stimuli. Hypercomplex cells are activated by edge,single-stopped(corner),double-stopped (tongue),slit(double-stopped)and dark bar(double-stopped).

\par In the visual cortex, perception cells are in the  order,  simple$\Longrightarrow$ complex$\Longrightarrow$ lower order hypercomplex $\Longrightarrow$ higher  order hypercomplex. Activation of  a lower stage is  influenced by  the position of the input patterns, and  higher stages are  position-invariant. There are several  contradictory to this structure, but no one   completely deny this hierarchical model.
\par Inspired by this work, Fukushima, K \cite{Stark1980} proposed a neural network model for pattern recognition called neocognitron.
In neocognitron , cells are arranged in  a number of  cascaded structure. Each  structure $U$ include a  simple cell layer $U_s$ and a complex cell layer $U_c$. This network is not affected by change in position or small distortion in the shape of patterns.  It is also capable of doing self-organization based on an unsupervised competitive learning algorithm\cite{Fukushima1982} in the first two layers and classification based on  supervised learning in the output layer.
\begin{figure}[ht]
 \centering
 \includegraphics[width=9cm]{Figures/neocognitron.eps}
\caption{Neocognitron\cite{Stark1980}}
\label{neo}
\end{figure}
\section{Convolutional Networks}
\par Neocognitron was improved by  Yann LeCun \cite{lecun-86},\cite{lecun-89d}, \cite{lecun-89e},\cite{lecun-90c}, \cite{lecun-90e} using backpropagation algorithm\cite{BRYSON1963} to train the entire system.
It uses  local receptive fields, share weights and  sub-sampling to achieve  shift, position and distortion invariance. A typical Convolutional Network called  LeNet-5 was proposed by   Yann LeCun et al.\cite{LeCun1998}  for document recognition. Using local receptive fields network can extract elementary visual features such as edges, end point, and corners. This features will combine to obtain high order features in the following layers.  Elementary feature detectors with identical weights can be useful in different parts of the image. So the units with  the same set of weights are arranged in plane and output from  the units of a plane is called  a feature map. Units in  a feature map perform the same operation on different parts of the same image. A convolution layer is composed of the set of feature maps with differently weighed units. In the implementation, a unit in the feature map scans the image and store the states in the feature map. This operation is equivalent to convolution with a kernel composed of a  set of weights and image.

\par
A typical convolutional  network composed of multiple stages with a filter bank layer, a non-linearity layer and a feature pooling layer \cite{lecun2010convolutional} followed by a classification network.\\
\begin{figure}[ht]
 \includegraphics[width=9.5cm]{Figures/convent.eps}
\caption{A typical ConvNet architecture \cite{lecun2010convolutional}}
\label{net}
\end{figure}
\emph{Filter Bank Layer:} This layer computes $y_{j}$ the convolution between a input feature map   $x_i$ and trainable filter kernel $k_{ij}$ 
. ie. $y_j=b_j+\sum_i {k_{ij}*x_i}$. Where $b_j$ is a trainable bias , $i$ and $j$ are  array indices and $*$ is the convolution operator.\\
\emph{Non-Linearity Layer:} This layer applies a non-linearity function such as $tanh(x)$ or $(1+e^{-x})^{-1}$ to unit output. But to reduce training time with gradient decent, new implementations uses the function $max(0,x)$. Units with this non-linearity is called Rectified Linear Units (ReLUs)\cite{Nair2010}.\\
\emph{Feature Pooling Layer:} It  reduces the dimension of  feature map by applying the techniques like averaging or max-pooling.
\begin{figure}[ht]
 \centering
 \includegraphics[width=8cm]{Figures/conv.eps}
\caption{Output of  a randomly initialized   convolution filter. Input image courtesy:dcsku.org}
\label{conv}
\end{figure}
But this architecture was limited by availability of the computing power and  sample data sets of that time. 
%2012

\section {Deep convolutional networks}
In last few years convolutional networks shows a significant performance improvement in many small scale image classification on data sets  such as MNIST\cite{Ciresan:2012g},CIFAR-10,CIFAR-100,SVHN\cite{lee2014deeply},STL-10 \cite{deepfwd}. One of the state-of-the-art error rate is from Krizhevsky, et al.\cite{Krizhevsky2012a}. It consist of 60 million parameters and 650,000 neurons, with five convolutional layers followed by max-pooling layers and three fully-connected layers. Data set used in this experiment was a subset of ImageNet dataset, used in the competition ImageNet Large-Scale Visual Recognition Challenge (ILSVRC)\cite{imagenet} and reported an error rate of 15.3\%. ILSVRC data set includes 1.2 million images that contain 1,000 categories. This network uses rectifier  as the function in neurons.  Even if it shows improvement in speed, this  ill-conditioned parameterization must be studied further to understand the effect in very large networks.

@@

\subsection{Network In Network }
Inspired by the work of Ian J. Goodfellow et al..\cite{Goodfellow2013} on max out networks,  Min Lin et al. \cite{Lin2013} introduced a micro-network in each convolution layer so that it will compute more abstract features.This network gave a state-of-the-art performance in  ILSVRC 2013 competition with an error rate of 12.95\%.They used NVIDIA TITAN GPU to train the network. Using multilayer perceptron instead of voting to approximate convex functions of each  local patches may result in a good accuracy, but this is equivalent to moving linear separability problem into another hyperspace. So if the input is in high frequency, this will end up modeling  large number of hidden layers and make the network structure more depend on the problem. It will diverge the idea  of the convolution network from  "learn anything" to a "local search".

@@

Matthew D. Zeiler and Rob Fergus\cite{Zeiler2013} presented a method to visualize the function of intermediate feature layers of convolutional networks and used as a diagnostic tool to improve  the model proposed by Krizhevsky et al. \cite{Krizhevsky2012a}. This method helped  them  to understand  the activation in the feature maps with respect to the input patterns. It shows Krizhevsky et al.'s architecture does not have  enough mid frequency coverage in the first layer filters and aliasing artifacts caused by  large stride in the first layer convolutions. Authors solved this problems by decreasing filter size to $7\times7$ and reducing stride to 2.This implementation won the  ILSVRC 2013 competition with an error rate of 11.74\%.Visualizations will help to debug the problem in some extend, But it become nearly impossible when there is large amount parameters involved. These  techniques may  more useful if it can relate to any of the learning formulations instead  vague  approximations of a non-invertible function.

@@

\subsection{Spatial pyramid pooling in Deep Convolutional Networks}
Instead of using fixed input size in convolutional networks, Kaiming He et al. \cite{He2014} suggested  the use of a spooling strategy called  spatial pyramid pooling(SPP)\cite{Grauman2005}\cite{1641019} to avoid cropping or warping of  images. It introduced a new layer on top of the convolution layer and perform aggregation  based on Bag-of-Words (BoW) model \cite{Sivic2003}. However, the classical backpropagation training methods expect layers to have a fixed   size.  This problem can be solved  using two fixed size networks with shared parameters and switched the network on alternate epochs. Network was trained using a single GeForce GTX Titan GPU with a starting  learning rate of 0.01 and achieved  a less  error rate of 8.06\% on  ILSVRC 2014 data set. 
\par 
This implementation improves the performance of baseline architectures including ZF-5\cite{Zeiler2013} , Convnet \cite{Krizhevsky2012a} and Overfeat-5/7 \cite{Sermanet2013}. Even if  the accuracy of convolutional networks will improve on multi-size training, multi-level pooling, and full-image representations, all this methods will increase both time and space complexity of the system.
%
